---
title: "behavior_analysis_notused"
format: html
---

Note: run behavior_cleaning and behavior_analysis first

## Behavior data

#### **Variance to mean ratio of urchin counts in a tank:**

A metric to assess the clumped vs. uniform vs. random distribution of urchins in a tank, and see how this varies when a star is/isn't present. This was calculated by finding the variance of urchin counts per pie slice in a tank and dividing this by the mean urchin count per pie slice in a tank averaged across behavioral assays in a trial.

Variance to mean of 1 = random, \>1 = clumped, \<1 = uniform

```{r}
#| layout-ncol: 2
#| label: raw-bhvr-variance

#VARIANCE TO MEAN RATIO of urchin distribution over treatments
#NOTE: v-to-mean of 1 = random, > 1 = clumped, < 1 = uniform
#should i find the variance of the var to mean ratios to find trial scale clustering dynamics?
ggplot(behavior_processed, 
       aes(x = Urch_habitat_treatment,
           y = trial_avg_var_to_mean,
           fill = Pred_treatment)) +
  
  geom_bar(stat = "identity", position = position_dodge()) +
  
  labs(title = "Trial avg variance-to-mean ratio of urchin slice counts",
       x = "Treatment",
       y = "Trial avg variance-to-mean ratio of urchin counts") +
  
  theme_classic()


ggplot(behavior_processed, 
       aes(x = Urch_habitat_treatment,
           y = trial_avg_var_to_mean,
           fill = Pred_treatment)) +
  
  geom_boxplot() +
  
  labs(title = "Variance to mean ratio of urchin slices across treatments",
       x = "Treatment",
       y = "Variance of percentage of urchins per slice") +
  
  theme_classic()
```

```{r}
#| layout-ncol: 2
#| label: mod-bhvr-variance

var2mean_mod <- glmmTMB::glmmTMB(var_to_mean ~ 
                         
                         Urch_habitat_treatment + 
                         Pred_treatment + 
                         Urch_habitat_treatment*Pred_treatment +
                         
                         (1| Trial/Tank),  
                       
                       data = behavior_processed,
                       
                       family = gaussian(link = "identity"))

sjPlot::tab_model(var2mean_mod)

predicted_var2mean_mod <- ggpredict(model = var2mean_mod, terms = c("Pred_treatment","Urch_habitat_treatment"))

ggplot(predicted_var2mean_mod, aes(group, predicted, fill = x)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), position=position_dodge()) +
  labs(title = "Predicted values of var to mean ratio from GLMM", x = "Urchin habitat", y = "Var-to-mean ratio", fill = "Pred treatment") +
  theme_classic() +
  scale_fill_manual(values = c('#DAF7A6','#5A5A5A') ) +
  ylab("Var-to-mean ratio") +
  labs(fill = "Predator treatment")
```

Due to overlapping error bars and non-significant effects of predator treatment in the model, there seems to be no clear effect of star presence on the clumped vs. uniform distribution of urchins in a tank. (Note: need to add error bars to bar chart).

#### **Crevice use**

The percentage of urchins in a tank in a crevice averaged over behavioral assays (at the trial scale).

```{r}
#| layout-ncol: 3
#| label: raw-bhvr-crev

ggplot(behavior_processed) +
  geom_histogram(aes(x = trial_avg_pcnt_in_crev, fill = Pred_treatment),position = "dodge") +
  facet_wrap(vars(Urch_habitat_treatment, Pred_treatment)) +
  labs(title = "Histogram for trial avg % of urchins in crevice") +
  theme_classic()

#bar charts of crevice use (looks same for behavior scale and trial scale)
ggplot(behavior_processed, 
       aes(x = Pred_treatment, y = trial_avg_pcnt_in_crev)) +
  
  geom_bar(aes(fill = Urch_habitat_treatment), position = position_dodge(), stat = "identity") +
  
  geom_errorbar(aes(ymin = trial_avg_pcnt_in_crev-trial_sd_pcnt_in_crev, 
                ymax = trial_avg_pcnt_in_crev+trial_sd_pcnt_in_crev,
                color = Urch_habitat_treatment),
                position = position_dodge()) +
  
  labs(title = "Bar chart of % of urchins in crevice") +
  theme_classic()

#boxplot of crevice use - looks diff for behavior scale
ggplot(behavior_processed) +
  geom_boxplot(aes(x = Pred_treatment, y = trial_avg_pcnt_in_crev, 
               fill = Urch_habitat_treatment)) +
  labs(title = "Box plot of % of urchins in crevice") +
  theme_classic()


```

```{r}
#| layout-ncol: 2
#| label: mod-bhvr-crev

crev_mod <- glmmTMB::glmmTMB(pcnt_in_crev ~ 
                         
                         Urch_habitat_treatment + 
                         Pred_treatment + 
                         Urch_habitat_treatment*Pred_treatment +
                         
                         (1| Trial/Tank),  
                         
                         data = behavior_processed,
                       
                       family = gaussian(link = "identity"))

sjPlot::tab_model(crev_mod)

predicted_crev_mod <- ggpredict(model = crev_mod, terms = c("Pred_treatment","Urch_habitat_treatment"))

ggplot(predicted_crev_mod, aes(group, predicted, fill = x)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), position=position_dodge()) +
  labs(title = "Predicted values of var to mean ratio from GLMM", x = "Urchin habitat", y = "Var-to-mean ratio", fill = "Pred treatment") +
  theme_classic() +
  scale_fill_manual(values = c('#DAF7A6','#5A5A5A') ) +
  ylab("Var-to-mean ratio") +
  labs(fill = "Predator treatment")
```

Again, due to overlapping error bars and non-significant fixed effects, there are no clear trends for crevice use by urchins dependent on source habitat/predator treatment.

#### Urchins associated with kelp

\% of urchins in a tank averaged over a trial where urchins are touching kelp (mouths over kelp) - raw data.

```{r}
#| layout-ncol: 2
#| label: raw-bhvr-onkelp

#boxplot of touching kelp
ggplot(behavior_processed) +
  geom_boxplot(aes(x = Urch_habitat_treatment, y = trial_avg_pcnt_on_kelp, 
               fill =  Pred_treatment)) +
  labs(title = "Box plot of % of urchins on kelp") +
  theme_classic()

```

#### Urchin spatial distribution in tank

Boxplots below show urchin distribution in tank at varying distances from the inflow. Urchins have a baseline tendency to be distributed near the inflow, which should be taken into account when looking at distribution when a star is/isn't present.

```{r}
#| label: raw-bhvr-urchdistribution

#boxplot of urchin distribution separated by treatment 
ggplot(behavior_processed) +
  geom_boxplot(aes(x = slice, y = urchin_slice_count, color = Treatment, fill = Treatment, alpha = 0.2)) +
  labs(title = "Urchin counts at different distances (in pie slices) from inflow",
       xlab = "Slices away from inflow",
       ylab = "Urchin slice count",
       caption = "Note: half slices are where crevices are located") +
  theme_classic()
```

```{r}
#| layout-ncol: 2
#| echo: false
#| message: false
#| label: mod-bhvr-urchdistribution

#currently, slices include half slices which are for crev, removing this for model
urch_count_non_crev <- behavior_processed %>%
  filter(slice %in% c(1,2,3,4,5,6,7,8))

urch_slice_mod <- glmmTMB::glmmTMB(urchin_slice_count ~ 
  
                        Urch_habitat_treatment + 
                        Pred_treatment + 
                        slice +
                        Urch_habitat_treatment*Pred_treatment*slice +
                        
                        (1| Trial/Tank),  
                      
                      data = urch_count_non_crev,
                      
                      family = gaussian(link = "identity"))

sjPlot::tab_model(urch_slice_mod)

predicted_urch_slice_mod <- ggpredict(model = urch_slice_mod, terms = c("Pred_treatment","Urch_habitat_treatment", "slice"))

#need to format this more - and also add urchin source hab treatment somehow.
ggplot(predicted_urch_slice_mod, aes(facet, predicted, fill = x)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), position=position_dodge()) 
```

## 
