---
title: "pycnopie_report_w_all_bhvr"
format: html
---

```{r}
###########################
###     behavior      ####
##########################

#note: slice = colored pie slice. position = outer/bottom/side/etc, aka orientation within slice. location = unique pairing of slice + position

#loading data:
behavior <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1yFkZm64m7c9b4EdSxK3-qR8B7I-5rWyk7CDs7fmmMKg/edit?gid=1070675104#gid=1070675104",
                                      "Behavior")



#Cleaning------------------------------------------------------------------------

#filtering behavior entries where total no of urch counted < 20
behavior1 <- behavior %>%
  filter(Total_num_urchin >= 20)

#changing crev count to doubles (used for pivot_longer)
behavior1$Num_urch_CREV1 <- as.double(behavior1$Num_urch_CREV1, na.rm = TRUE)
behavior1$Num_urch_CREV2 <- as.double(behavior1$Num_urch_CREV2, na.rm = TRUE)
behavior1$Num_urch_CREV3 <- as.double(behavior1$Num_urch_CREV3, na.rm = TRUE)
behavior1$Num_urch_CREV4 <- as.double(behavior1$Num_urch_CREV4, na.rm = TRUE)
behavior1$Num_urch_CREV5 <- as.double(behavior1$Num_urch_CREV5, na.rm = TRUE)


behavior_cleaned <- behavior1 %>%
  pivot_longer(Green_outer_SIDE:Num_urch_CREV5, names_to = "urchin_location", values_to = "urchin_location_count") %>% 
  
  #changing crev locations to pie slice location 
  mutate(crev_count = case_when(grepl("CREV",urchin_location) ~ urchin_location_count, .default = 0),
         #renaming values in urchin_location column to be more descriptive
         urchin_location = case_when(
           urchin_location == "Num_urch_CREV1" ~ ".5_Outer_bottom",
           urchin_location == "Num_urch_CREV2" ~ "2.5_Outer_bottom",
           urchin_location == "Num_urch_CREV3" ~ "4.5_Outer_bottom",
           urchin_location == "Num_urch_CREV4" ~ "6.5_Outer_bottom",
           urchin_location == "Num_urch_CREV5" ~ "0_Outer_bottom",
           urchin_location == "Green_outer_SIDE" ~ "1_Outer_side",
           urchin_location == "Green_outer_BOTTOM" ~ "1_Outer_bottom",
           urchin_location == "Green_INNER" ~ "1_Inner",
           urchin_location == "Red_outer_SIDE" ~ "2_Outer_side",
           urchin_location == "Red_outer_BOTTOM" ~ "2_Outer_bottom",
           urchin_location == "Red_INNER" ~ "2_Inner",
           urchin_location == "Yellow_outer_SIDE" ~ "3_Outer_side",
           urchin_location == "Yellow_outer_BOTTOM" ~ "3_Outer_bottom",
           urchin_location == "Yellow_INNER" ~ "3_Inner",
           urchin_location == "Blue_outer_SIDE" ~ "4_Outer_side",
           urchin_location == "Blue_outer_BOTTOM" ~ "4_Outer_bottom",
           urchin_location == "Blue_INNER" ~ "4_Inner",
           urchin_location == "White_outer_SIDE" ~ "6_Outer_side", #are you sure
           urchin_location == "White_outer_BOTTOM" ~ "6_Outer_bottom",
           urchin_location == "White_INNER" ~ "6_Inner",
           urchin_location == "Orange_outer_SIDE" ~ "5_Outer_side",
           urchin_location == "Orange_outer_BOTTOM" ~ "5_Outer_bottom",
           urchin_location == "Orange_INNER" ~ "5_Inner",
           urchin_location == "Black_outer_SIDE" ~ "7_Outer_side",
           urchin_location == "Black_outer_BOTTOM" ~ "7_Outer_bottom",
           urchin_location == "Black_INNER" ~ "7_Inner",
           urchin_location == "White2_outer_SIDE" ~ "8_Outer_side",
           urchin_location == "White2_outer_BOTTOM" ~ "8_Outer_bottom",
           urchin_location == "White2_INNER" ~ "8_Inner",
             
             .default = urchin_location)) %>%
  
  #making a column for slice and a column for position
  separate(urchin_location, c("urchin_slice", "urchin_position"), "_",extra = "merge") %>%

  separate(`Pycno1 position`, c("pycno1_position","pycno1_slice"), "\\_(?!.*_)", extra = "merge") %>%
  separate(`Pycno2 position`, c("pycno2_position","pycno2_slice"), "\\_(?!.*_)", extra = "merge") %>%
  
  
  mutate(pycno1_slice = case_when(pycno1_slice == "green" ~ 1,
                                  pycno1_slice == "red" ~ 2,
                                  pycno1_slice == "yellow" ~ 3,
                                  pycno1_slice == "blue" ~ 4,
                                  pycno1_slice == "orange" ~ 5,
                                  pycno1_slice == "white" ~ 6,
                                  pycno1_slice == "black" ~ 7,
                                  pycno1_slice == "white2" ~ 8,
                                  pycno1_slice == "crev" ~ 0),
         
         pycno2_slice = case_when(pycno2_slice == "green" ~ 1,
                                  pycno2_slice == "red" ~ 2,
                                  pycno2_slice == "yellow" ~ 3,
                                  pycno2_slice == "blue" ~ 4,
                                  pycno2_slice == "orange" ~ 5,
                                  pycno2_slice == "white" ~ 6,
                                  pycno2_slice == "black" ~ 7,
                                  pycno2_slice == "white2" ~ 8,
                                  pycno2_slice == "crev" ~ 0)) %>%
  
  mutate(urchin_slice = as.numeric(urchin_slice)) %>%
  mutate(urchin_location_count = case_when(is.na(urchin_location_count) ~ 0, .default = urchin_location_count)) %>% #for now - changed urchin count NA to zero - but think about whether or not this is ok!!!!
  
#averaging over position scale to get one value per slice
  group_by(Tank, Treatment, Date, Day_numrecord, Trial, urchin_slice) %>%
  mutate(urchin_slice_count = sum(urchin_location_count)) %>%
  ungroup() %>%
  
  separate(Treatment, "_(?=[^_]*$)",
           into = c("Pred_treatment", "Urch_habitat_treatment"), 
           remove = FALSE) 

  



#still need to figure this part out.
  #mutate(pyc_slice_count = case_when(pycno1_slice != 0 | urchin_slice == pycno2_slice ~ 1, #NOTE: this is number of pycnos in a SLICE - not per position (outer/bottom/etc) - can add this variable too if you want to compare it later!
                                     #slice == pycno1_slice & slice == pycno2_slice ~ 2, #for some reason no 2 vvalues showing up - in the data or prob w code?
                                    # is.na(pycno1_slice) & Treatment %in% c("Star_KF", "Star_B") ~ NA, #turns pycno_count to 0 if pyc slice wasn't recorded in a treatment w a star
                                   #  is.na(pycno2_slice) & Treatment %in% c("Star_KF", "Star_B") ~ NA,
                                     #.default = 0))

```

```{r}
behavior_processed <- behavior_cleaned %>%
  
  #removing position for now - might want to change later - changing to scale of day/num record
  select(-urchin_position, -urchin_location_count) %>%
  unique() %>% 
  
  
  #adding distance to pycno
  mutate(dist_slice_pyc1 = abs(urchin_slice - pycno1_slice),
         dist_slice_pyc2 = abs(urchin_slice - pycno2_slice)) %>%
  

  #adding variance of pie slices + crevice use at day/numrecord scale. currently, num urch on kelp is at the 
  group_by(Trial, Treatment, Day_numrecord) %>%
  mutate(variance_of_urch_distrib = var(urchin_slice_count, na.rm = TRUE), #one value per tank
         var_to_mean = variance_of_urch_distrib/mean(urchin_slice_count),
         pcnt_in_crev = sum(crev_count, na.rm=TRUE)/Total_num_urchin *100,
         pcnt_on_kelp = mean(Num_urch_on_kelp, na.rm = TRUE)/Total_num_urchin *100) %>%
  ungroup() %>%
  
  #averaged between time points (trial scale)
  group_by(Trial, Treatment) %>%
  mutate(trial_avg_pcnt_in_crev = mean(pcnt_in_crev),
         trial_sd_pcnt_in_crev = sd(pcnt_in_crev, na.rm = TRUE),
         trial_avg_var_to_mean = mean(var_to_mean),
         trial_avg_pcnt_on_kelp = mean(pcnt_on_kelp),
         trial_sd_pcnt_on_kelp = sd(pcnt_on_kelp)) %>%
  ungroup()

#this is for a appending avg pycno pos to kelp data
behavior_processed_forkelp <- behavior_processed %>%
  group_by(Trial, Treatment) %>% #going from behavioral assay scale to trial scale
  summarise(avg_pycno1_slice = mean(pycno1_slice, na.rm = TRUE),
            avg_pycno2_slice = mean(pycno2_slice, na.rm = TRUE),
            #closer to 1 = upstream
            up_dn_pyc1_ratio = sum(up_dn_pyc1=="upstream", na.rm = TRUE)/sum(up_dn_pyc1=="downstream", na.rm = TRUE),
            up_dn_pyc2_ratio = sum(up_dn_pyc2=="upstream")/sum(up_dn_pyc2=="downstream")) %>%
  ungroup()

```

## Behavior data

Other things I did not included here: plotted variance to mean ratio of urchins (way to see if distribution is random, clumped, uniform, etc)

Variance to mean of 1 = random, \>1 = clumped, \<1 = uniform

```{r}
#| layout-ncol: 2
#| label: mod-bhvr-variance

var2mean_mod <- glmmTMB::glmmTMB(var_to_mean ~ 
                         
                         Urch_habitat_treatment + 
                         Pred_treatment + 
                         Urch_habitat_treatment*Pred_treatment +
                         
                         (1| Trial/Tank),  
                       
                       data = behavior_processed,
                       
                       family = gaussian(link = "identity"))

sjPlot::tab_model(var2mean_mod)

predicted_var2mean_mod <- ggpredict(model = var2mean_mod, terms = c("Pred_treatment","Urch_habitat_treatment"))

ggplot(predicted_var2mean_mod, aes(group, predicted, fill = x)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), position=position_dodge()) +
  labs(title = "Predicted values of var to mean ratio from GLMM", x = "Urchin habitat", y = "Var-to-mean ratio", fill = "Pred treatment") +
  theme_classic() +
  scale_fill_manual(values = c('#DAF7A6','#5A5A5A') ) +
  ylab("Var-to-mean ratio") +
  labs(fill = "Predator treatment")
```

Due to overlapping error bars and non-significant effects of predator treatment in the model, there seems to be no clear effect of star presence on the clumped vs. uniform distribution of urchins in a tank. (Note: need to add error bars to bar chart).

#### **Crevice use**

The percentage of urchins in a tank in a crevice averaged over behavioral assays (at the trial scale).

```{r}
#| layout-ncol: 3
#| label: raw-bhvr-crev

ggplot(behavior_processed) +
  geom_histogram(aes(x = trial_avg_pcnt_in_crev, fill = Pred_treatment),position = "dodge") +
  facet_wrap(vars(Urch_habitat_treatment, Pred_treatment)) +
  labs(title = "Histogram for trial avg % of urchins in crevice") +
  theme_classic()

#bar charts of crevice use (looks same for behavior scale and trial scale)
ggplot(behavior_processed, 
       aes(x = Pred_treatment, y = trial_avg_pcnt_in_crev)) +
  
  geom_bar(aes(fill = Urch_habitat_treatment), position = position_dodge(), stat = "identity") +
  
  geom_errorbar(aes(ymin = trial_avg_pcnt_in_crev-trial_sd_pcnt_in_crev, 
                ymax = trial_avg_pcnt_in_crev+trial_sd_pcnt_in_crev,
                color = Urch_habitat_treatment),
                position = position_dodge()) +
  
  labs(title = "Bar chart of % of urchins in crevice") +
  theme_classic()

#boxplot of crevice use - looks diff for behavior scale
ggplot(behavior_processed) +
  geom_boxplot(aes(x = Pred_treatment, y = trial_avg_pcnt_in_crev, 
               fill = Urch_habitat_treatment)) +
  labs(title = "Box plot of % of urchins in crevice") +
  theme_classic()


```

```{r}
#| layout-ncol: 2
#| label: mod-bhvr-crev

crev_mod <- glmmTMB::glmmTMB(pcnt_in_crev ~ 
                         
                         Urch_habitat_treatment + 
                         Pred_treatment + 
                         Urch_habitat_treatment*Pred_treatment +
                         
                         (1| Trial/Tank),  
                         
                         data = behavior_processed,
                       
                       family = gaussian(link = "identity"))

sjPlot::tab_model(crev_mod)

predicted_crev_mod <- ggpredict(model = crev_mod, terms = c("Pred_treatment","Urch_habitat_treatment"))

ggplot(predicted_crev_mod, aes(group, predicted, fill = x)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), position=position_dodge()) +
  labs(title = "Predicted values of var to mean ratio from GLMM", x = "Urchin habitat", y = "Var-to-mean ratio", fill = "Pred treatment") +
  theme_classic() +
  scale_fill_manual(values = c('#DAF7A6','#5A5A5A') ) +
  ylab("Var-to-mean ratio") +
  labs(fill = "Predator treatment")
```

Again, due to overlapping error bars and non-significant fixed effects, there are no clear trends for crevice use by urchins dependent on source habitat/predator treatment.

#### Urchins associated with kelp

\% of urchins in a tank averaged over a trial where urchins are touching kelp (mouths over kelp) - raw data.

```{r}
#| layout-ncol: 2
#| label: raw-bhvr-onkelp

#boxplot of touching kelp
ggplot(behavior_processed) +
  geom_boxplot(aes(x = Urch_habitat_treatment, y = trial_avg_pcnt_on_kelp, 
               fill =  Pred_treatment)) +
  labs(title = "Box plot of % of urchins on kelp") +
  theme_classic()

```

#### Urchin spatial distribution in tank

Boxplots below show urchin distribution in tank at varying distances from the inflow. Urchins have a baseline tendency to be distributed near the inflow, which should be taken into account when looking at distribution when a star is/isn't present.

```{r}
#| label: raw-bhvr-urchdistribution

#boxplot of urchin distribution separated by treatment 
ggplot(behavior_processed) +
  geom_boxplot(aes(x = slice, y = urchin_slice_count, color = Treatment, fill = Treatment, alpha = 0.2)) +
  labs(title = "Urchin counts at different distances (in pie slices) from inflow",
       xlab = "Slices away from inflow",
       ylab = "Urchin slice count",
       caption = "Note: half slices are where crevices are located") +
  theme_classic()
```

```{r}
#| layout-ncol: 2
#| echo: false
#| message: false
#| label: mod-bhvr-urchdistribution

#currently, slices include half slices which are for crev, removing this for model
urch_count_non_crev <- behavior_processed %>%
  filter(slice %in% c(1,2,3,4,5,6,7,8))

urch_slice_mod <- glmmTMB::glmmTMB(urchin_slice_count ~ 
  
                        Urch_habitat_treatment + 
                        Pred_treatment + 
                        slice +
                        Urch_habitat_treatment*Pred_treatment*slice +
                        
                        (1| Trial/Tank),  
                      
                      data = urch_count_non_crev,
                      
                      family = gaussian(link = "identity"))

sjPlot::tab_model(urch_slice_mod)

predicted_urch_slice_mod <- ggpredict(model = urch_slice_mod, terms = c("Pred_treatment","Urch_habitat_treatment", "slice"))

#need to format this more - and also add urchin source hab treatment somehow.
ggplot(predicted_urch_slice_mod, aes(facet, predicted, fill = x)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), position=position_dodge()) 
```

```{r}
#calculation of predation rate: 
#total number of urchins eaten / number of urchins in star treatments:
#this was calculated manually based on notes written of when urchins eaten

#% of urchins consumed
22/(24*24)*100 #22 total urchins consumed, 24 total star treatments, 24 urchins per tank

#predation rate: of two stars, so divided by 2
#pred1 - 1 urchin eaten by 2 stars (48 hr trial duration)
p1 <- (1/48)/2 
p2 <- (1/48)/2
p3 <- (2/48)/2
p4 <- (1/48)/2
p5 <- (1/48)/2
p6 <- (2/48)/2
p7 <- (1/48)/2
p8 <- (4/48)/2
p9 <- (1/48)/2
p10 <- (1/48)/2
p11 <- (1/48)/2
p12 <- (2/48)/2
p13 <- (1/48)/2
p14 <- (1/48)/2
p15 <- (2/48)/2

#mean pred rate
mean(c(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12,p13,p14,p15))
```
