---
title: "PycnoPie"
format: 
  html:
    code-fold: true
    page-layout: full
editor: visual
date: last-modified
message: false
---

**Analysis of urchin grazing data for PycnoPie experiment on sunflower star indirect effects on kelp with urchins from different source habitats conducted in Sitka, AK in summer of 2024.**

```{r}
#| echo: false
#| warning: false

library(here)
library(tidyverse)
library(googlesheets4)
library(strex)
library(lme4)
library(performance)
library(sjPlot)
library(glmmTMB)
library(DHARMa)
library(ggeffects)
library(effects)
library(car)
library(emmeans)
library(ggsignif)
library(flextable)
library(lmerTest)
library(ggpubr)

kelp <- read.csv("data/raw/exp2_kelp_weight.csv")
behavior <- read.csv("data/raw/exp2_behavior.csv")


source(here::here("./scripts/00_cleaning.R"))
source(here::here("./scripts/01_manipulation.R"))


```

```{r}
#| echo: false

#default formatting for all tables in this report
set_flextable_defaults(
  font.size = 11, font.family = "Arial",
  font.color = "#333333",
  table.layout = "fixed",
  border.color = "gray",
  padding.top = 3, padding.bottom = 3,
  padding.left = 4, padding.right = 4)
```

```{r}
#| echo: false
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Kelp grazing data

### **Building** models and contrasts

Investigating distribution of response variable and test for normality.

```{r}
#| layout-ncol: 2
#| label: model-kelp
#| results: hide



#releveling reference groups 
kelp_processed$Urch_habitat_treatment <- relevel(kelp_processed$Urch_habitat_treatment, ref = "KF")
kelp_processed$Pred_treatment <- relevel(kelp_processed$Pred_treatment, ref = "Control")


#investigating distributions 
ggplot(kelp_processed, aes(x = weight_diff)) +
  geom_density() +
  facet_wrap(vars(Treatment)) +
  theme_bw()


#test for normality
qqPlot(kelp_processed$weight_diff)

```

Linear mixed-effects model for difference in kelp weight

```{r}
#| results: hide

#building model

model <- lmer(weight_diff ~ 
                  
              Pred_treatment + Urch_habitat_treatment + 
              Pred_treatment*Urch_habitat_treatment + (1|Trial/Tank),
                
              data = kelp_processed)

summary(model)

#anova formatted in a flextable
anova(model) %>%
  data.frame() %>%
  as.tibble(rownames = " ", colnames = " ") %>%
   mutate(" " = case_when(row_number() == 1 ~ "Predator treatment",
                        row_number() == 2 ~"Urchin source habitat",
                        row_number() == 3 ~"Predator treatment:Urchin source habitat"),
          Pr..F. = case_when(Pr..F. < 0.0001 ~ "< 0.0001",
                             row_number() == 1 ~ "0.0003",
                             .default= sprintf("%.3f",Pr..F.))) %>%

  flextable() %>%
  #header formatting
  set_header_labels(Df = 'd.f.',
                    Sum.Sq = "sum sq",
                    Mean.Sq = "mean sq",
                    F.value = "f-value",
                    Pr..F. = "p-value") %>%
  bold(i = 1, part = "header") %>%
  align(align = 'center', part = 'header') %>%
  #body formatting
  colformat_double(j = c(2:6), digits = 3) %>%
  bold(~Pr..F. < 0.05, j = "Pr..F.") %>%
  vline_left() %>%
  vline_right() %>%
  
  autofit()  
  
```

### **Post-hoc tests**

Calculating estimated marginal means from the linear mixed-effects model, and calculating pairwise contrasts based on a priori hypotheses based on custom contrast weights

```{r}

#estimated marginal means
EMM <- emmeans(model, ~ Pred_treatment + Urch_habitat_treatment + Pred_treatment*Urch_habitat_treatment)
EMM

#specifying weights for contrasts
contrast_weights <- list(
  "a) Control kelp forest vs. Control barren" = c(control_kf = 1,
                                             star_kf = 0,
                                             control_b = -1,
                                             star_b = 0),
  
  "b) Star kelp forest vs. Star barren" = c(control_kf = 0,
                                       star_kf = 1,
                                       control_b = 0,
                                       star_b = -1),
  
  "c) Control kelp forest vs. Star kelp forest" = c(control_kf = 1,
                                               star_kf = -1,
                                               control_b = 0,
                                               star_b = 0),
  
  "d) Control barren vs. Star barren" = c(control_kf = 0,
                                       star_kf = 0,
                                       control_b = 1,
                                       star_b = -1)
  
)

#pairwise contrasts w above weighting scheme
mod_contrasts <- contrast(EMM, contrast_weights)

#flextable formatting
ft <- mod_contrasts %>% 
  
  data.frame() %>%  
  mutate(p.value = ifelse(p.value < 0.0001, "< 0.0001",
                          round(p.value,3))) %>%
  #hypothesis column
   mutate("Hypothesis" = c("Urchins from kelp forests will graze less than urchins from barrens",
                          "Urchins in the presence of a star will graze less when they are from kelp forests compared to urchins from barrens",
                          "Urchins from kelp forests will graze more when they are not around a star",
                          "Urchins from barrens will graze more when they are not around a star"),.after = "contrast") %>% 
  
  flextable()  %>%
  #header formatting
  set_header_labels(contrast = 'Contrast',
                    emmean = 'Estimate',
                    SE = 's.e.',
                    df = 'd.f.',
                    t.ratio = "t-ratio",
                    p.value = "p-value") %>%
  bold(i = 1, part = "header") %>%
  italic(part = "all", j = 1) %>%

  align(align = 'center', part = 'header') %>%

  #body formatting
  #change size of hypothesis column 
  fontsize(j = 2, size = 8, part = "body") %>%
  colformat_double(j = c(3:7), digits = 2) %>%
  bold(~ p.value < 0.05, j = "p.value") %>%
  vline_left() %>%
  vline_right() %>%
  
  autofit()  

ft

  

```

### Plots:

Bar chart of raw data showing change in kelp weight across treatments. Significance bars manually added based on significant pairwise contrasts.

```{r}
#| results: hide

#plots of raw data:
#releveling reference groups (this is just bc i manually labelled x axis formatting stuff based on indexing position, and first is barren second is kf)
kelp_processed$Urch_habitat_treatment <- relevel(kelp_processed$Urch_habitat_treatment, ref = "B")
kelp_processed$Pred_treatment <- relevel(kelp_processed$Pred_treatment, ref = "Control")

#bar chart 

dw <- 0.9 # dodge width
off <- dw / 4

ggplot(kelp_processed, 
       aes(x = Urch_habitat_treatment, y = weight_diff_mean, 
           fill = Pred_treatment)) + 
  
  geom_bar(stat = "identity", 
           position = position_dodge(width=dw)) +
  
  geom_errorbar(aes(ymin = weight_diff_mean-weight_diff_se, 
                    ymax =weight_diff_mean+weight_diff_se),
                position= position_dodge(width=dw),
                width = 0.1) +
  
   labs(title = "Kelp grazed by urchins from different source habitats", 
        x = "Urchin habitat", y = "Mean difference in kelp weight (g)", #specify across the 12 blades in a tank
        fill = "Pred treatment",
        caption = "* p < 0.05, ** p < 0.01, *** p < 0.0001 ") +
 
  theme_classic() +
  theme(plot.caption = element_text(color = "grey40", hjust = 1.5)) +
  
  scale_fill_manual(values = c('#5A5A5A','#c994c7')) +
  ylab("Mean difference in kelp weight (g)") +
  scale_x_discrete(labels = c("Barren","Kelp forest")) +
  labs(fill = "Predator treatment") +
  
  #significance bars:
  #Barren control vs. barren star (***)
  geom_bracket(
    inherit.aes = FALSE,
    xmin = 1 - off,  # barren control
    xmax = 1 + off,  # barren star
    y.position = 9.9,
    label = "***",
    color = "grey40", alpha = 0.7, size = 0.5, vjust = 0.5
  ) +
  #Barren control vs. kelp forest control (***)
  geom_bracket(
    inherit.aes = FALSE,
    xmin = 1 - off,  # barren control
    xmax = 2 - off,  # kelp forest control
    y.position = 11,
    label = "***",
    color = "grey40", alpha = 0.7, size = 0.5, vjust = 0.5
  ) +
  #Barren star vs. kelp forest star (***)
  geom_bracket(
    inherit.aes = FALSE,
    xmin = 1 + off,  # barren star
    xmax = 2 + off,  # kelp forest star
    y.position = 5.3,
    label = "***",
    color = "grey40", alpha = 0.7, size = 0.5, vjust = 0.5
  ) 
  



```

## Key takeaways:

> Urchins graze less around a star for both kelp forest and barren treatments
>
> Grazing rates for urchins from a barren are more affected by the presence of a star than urchins from a kelp forest
