---
title: "PycnoPie"
format: 
  html:
    code-fold: true
    page-layout: full
editor: visual
date: last-modified
---

*Select results and interpretation from PycnoPie experiment run in the summer of 2024 on indirect effects of Pycnopodia helianthoides on grazing rates of red urchins from kelp forests and urchin barrens. Cleaning and processing scripts can be found [here](https://github.com/nikita-sridhar/PycnoPie/tree/main/scripts).*

```{r}
#| echo: false

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
#| message: false
#| echo: false

#loading libraries
library(tidyverse)
library(here)
library(googlesheets4)
library(strex)
library(lme4)
library(performance)
library(sjPlot)
library(glmmTMB)
library(DHARMa)
library(ggeffects)
library(effects)
```

```{r}
#| message: false
#| echo: false

#source(here::here("./scripts/00_cleaning.R"))
#source(here::here("./scripts/01_manipulation.R"))

kelp_processed <- read_csv(here::here("data/manipulated/processed_kelp/kelp_processed.csv"))
behavior_processed <- read_csv(here::here("data/manipulated/processed_behavior/behavior_processed.csv"))

```

## Kelp grazing data

#### **Raw data:** % change in kelp grazed by urchins

```{r}
#| layout-ncol: 2
#| label: raw-kelp

#kelp grazed by predator/urchin treatment
ggplot(data = subset(kelp_processed, !is.na(pcnt_grazed)), aes(x = Urch_habitat_treatment, y = pcnt_grazed, fill = Pred_treatment) ) +
  geom_boxplot() +
  theme_classic() +
  scale_fill_manual(values = c('#5A5A5A','#DAF7A6') ) +
  ylab("% Change Kelp Grazed") +
  xlab("Urchin habitat ") +
  labs(fill = "Predator treatment")

```

#### **Model data:** General linear mixed model for % change in kelp grazed by urchins

```{r}
#| layout-ncol: 2
#| label: model-kelp

#building model
kelp_mod <- glmmTMB::glmmTMB(pcnt_grazed ~ 
                         
                         Urch_habitat_treatment + 
                         Pred_treatment + 
                         Urch_habitat_treatment*Pred_treatment +
                         
                         (1| Trial/Tank/Kelp_ID),  
                       
                       data = kelp_processed,
                       
                       family = gaussian(link = "identity"))

sjPlot::tab_model(kelp_mod)

predicted_kelp_mod <- ggpredict(model = kelp_mod, terms = c("Pred_treatment","Urch_habitat_treatment"))

#calculating SD by group for raw data for error bars
kelp_processed_sd <- kelp_processed %>%
  group_by(Urch_habitat_treatment, Pred_treatment) %>%
  mutate(sd = sd(pcnt_grazed)) %>%
  ungroup()

#bar chart of model estimates with error bars from raw data's standard error
ggplot(predicted_kelp_mod, 
       aes(group, predicted, fill = x)) +
  
  geom_bar(stat = "identity", 
           position = position_dodge()) +
  
  geom_errorbar(data = kelp_processed_sd,
    aes(ymin = mean(pcnt_grazed)-sd, ymax = mean(pcnt_grazed)+sd), 
                position=position_dodge()) +
 
   labs(title = "Predicted values of % kelp grazed from GLMM", 
        x = "Urchin habitat", y = "% kelp grazed", 
        fill = "Pred treatment") +
 
  theme_classic() +
  scale_fill_manual(values = c('#DAF7A6','#5A5A5A') ) +
  ylab("% Change Kelp Grazed") +
  labs(fill = "Predator treatment")

#overlay of mean model estimate with raw data - defeats purpose of visualizing model estimate
ggplot() +
  
  #model plot (mean GLMM value)
    geom_point(data = predicted_kelp_mod, 
               aes(x = group, y = predicted, color = x), 
               position = position_dodge(0.5)) +
  
  geom_errorbar(data = kelp_processed, 
                aes(x = Urch_habitat_treatment, ymin = conf.low, ymax = conf.high, color = x), 
                position = position_dodge(0.5), 
                width = .1) +

  #overlayed raw data
  geom_point(data = kelp_processed, 
             aes(x = Urch_habitat_treatment, y = pcnt_grazed, color = Pred_treatment), 
             position = position_dodge(0.5), 
             alpha = 0.1) +
  
  labs(title = "Predicted values of % kelp grazed from GLMM overlayed with raw data", x = "Urchin habitat", y = "% kelp grazed", fill = "Pred treatment") +
  theme_classic() +
  ylab("% Change Kelp Grazed") +
  labs(fill = "Predator treatment")


```

The fixed effects of urchin habitat, predator treatment, and the interaction term are all significant in the GLMM for % kelp grazed. The plot of predicted values from the GLMM shows clearer trends in grazing than the boxplot of raw data. Both kelp forest and barren urchins reduce grazing around a star. Urchins from a barren reduce grazing more than urchins from a kelp forest. The reduction of grazing of kelp forest urchins around a star may be enough to allow for net kelp growth (note negative bar - this may be due to mucus production or growth; regardless, grazing suppression in this treatment was sufficient to allow for positive increase in kelp mass).

## Behavior data

#### **Variance to mean ratio of urchin counts in a tank:**

A metric to assess the clumped vs. uniform vs. random distribution of urchins in a tank, and see how this varies when a star is/isn't present. This was calculated by finding the variance of urchin counts per pie slice in a tank and dividing this by the mean urchin count per pie slice in a tank averaged across behavioral assays in a trial.

Variance to mean of 1 = random, \>1 = clumped, \<1 = uniform

```{r}
#| layout-ncol: 2
#| label: raw-bhvr-variance

#VARIANCE TO MEAN RATIO of urchin distribution over treatments
#NOTE: v-to-mean of 1 = random, > 1 = clumped, < 1 = uniform
#should i find the variance of the var to mean ratios to find trial scale clustering dynamics?
ggplot(behavior_processed, 
       aes(x = Urch_habitat_treatment,
           y = trial_avg_var_to_mean,
           fill = Pred_treatment)) +
  
  geom_bar(stat = "identity", position = position_dodge()) +
  
  labs(title = "Trial avg variance-to-mean ratio of urchin slice counts",
       x = "Treatment",
       y = "Trial avg variance-to-mean ratio of urchin counts") +
  
  theme_classic()


ggplot(behavior_processed, 
       aes(x = Urch_habitat_treatment,
           y = trial_avg_var_to_mean,
           fill = Pred_treatment)) +
  
  geom_boxplot() +
  
  labs(title = "Variance to mean ratio of urchin slices across treatments",
       x = "Treatment",
       y = "Variance of percentage of urchins per slice") +
  
  theme_classic()
```

```{r}
#| layout-ncol: 2
#| label: mod-bhvr-variance

var2mean_mod <- glmmTMB::glmmTMB(var_to_mean ~ 
                         
                         Urch_habitat_treatment + 
                         Pred_treatment + 
                         Urch_habitat_treatment*Pred_treatment +
                         
                         (1| Trial/Tank),  
                       
                       data = behavior_processed,
                       
                       family = gaussian(link = "identity"))

sjPlot::tab_model(var2mean_mod)

predicted_var2mean_mod <- ggpredict(model = var2mean_mod, terms = c("Pred_treatment","Urch_habitat_treatment"))

ggplot(predicted_var2mean_mod, aes(group, predicted, fill = x)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), position=position_dodge()) +
  labs(title = "Predicted values of var to mean ratio from GLMM", x = "Urchin habitat", y = "Var-to-mean ratio", fill = "Pred treatment") +
  theme_classic() +
  scale_fill_manual(values = c('#DAF7A6','#5A5A5A') ) +
  ylab("Var-to-mean ratio") +
  labs(fill = "Predator treatment")
```

Due to overlapping error bars and non-significant effects of predator treatment in the model, there seems to be no clear effect of star presence on the clumped vs. uniform distribution of urchins in a tank. (Note: need to add error bars to bar chart).

#### **Crevice use**

The percentage of urchins in a tank in a crevice averaged over behavioral assays (at the trial scale).

```{r}
#| layout-ncol: 3
#| label: raw-bhvr-crev

ggplot(behavior_processed) +
  geom_histogram(aes(x = trial_avg_pcnt_in_crev, fill = Pred_treatment),position = "dodge") +
  facet_wrap(vars(Urch_habitat_treatment, Pred_treatment)) +
  labs(title = "Histogram for trial avg % of urchins in crevice") +
  theme_classic()

#bar charts of crevice use (looks same for behavior scale and trial scale)
ggplot(behavior_processed, 
       aes(x = Pred_treatment, y = trial_avg_pcnt_in_crev)) +
  
  geom_bar(aes(fill = Urch_habitat_treatment), position = position_dodge(), stat = "identity") +
  
  geom_errorbar(aes(ymin = trial_avg_pcnt_in_crev-trial_sd_pcnt_in_crev, 
                ymax = trial_avg_pcnt_in_crev+trial_sd_pcnt_in_crev,
                color = Urch_habitat_treatment),
                position = position_dodge()) +
  
  labs(title = "Bar chart of % of urchins in crevice") +
  theme_classic()

#boxplot of crevice use - looks diff for behavior scale
ggplot(behavior_processed) +
  geom_boxplot(aes(x = Pred_treatment, y = trial_avg_pcnt_in_crev, 
               fill = Urch_habitat_treatment)) +
  labs(title = "Box plot of % of urchins in crevice") +
  theme_classic()


```

```{r}
#| layout-ncol: 2
#| label: mod-bhvr-crev

crev_mod <- glmmTMB::glmmTMB(pcnt_in_crev ~ 
                         
                         Urch_habitat_treatment + 
                         Pred_treatment + 
                         Urch_habitat_treatment*Pred_treatment +
                         
                         (1| Trial/Tank),  
                         
                         data = behavior_processed,
                       
                       family = gaussian(link = "identity"))

sjPlot::tab_model(crev_mod)

predicted_crev_mod <- ggpredict(model = crev_mod, terms = c("Pred_treatment","Urch_habitat_treatment"))

ggplot(predicted_crev_mod, aes(group, predicted, fill = x)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), position=position_dodge()) +
  labs(title = "Predicted values of var to mean ratio from GLMM", x = "Urchin habitat", y = "Var-to-mean ratio", fill = "Pred treatment") +
  theme_classic() +
  scale_fill_manual(values = c('#DAF7A6','#5A5A5A') ) +
  ylab("Var-to-mean ratio") +
  labs(fill = "Predator treatment")
```

Again, due to overlapping error bars and non-significant fixed effects, there are no clear trends for crevice use by urchins dependent on source habitat/predator treatment.

#### Urchins associated with kelp

\% of urchins in a tank averaged over a trial where urchins are touching kelp (mouths over kelp) - raw data.

```{r}
#| layout-ncol: 2
#| label: raw-bhvr-onkelp

#boxplot of touching kelp
ggplot(behavior_processed) +
  geom_boxplot(aes(x = Urch_habitat_treatment, y = trial_avg_pcnt_on_kelp, 
               fill =  Pred_treatment)) +
  labs(title = "Box plot of % of urchins on kelp") +
  theme_classic()

```

#### Urchin spatial distribution in tank

Boxplots below show urchin distribution in tank at varying distances from the inflow. Urchins have a baseline tendency to be distributed near the inflow, which should be taken into account when looking at distribution when a star is/isn't present.

```{r}
#| label: raw-bhvr-urchdistribution

#boxplot of urchin distribution separated by treatment 
ggplot(behavior_processed) +
  geom_boxplot(aes(x = slice, y = urchin_slice_count, color = Treatment, fill = Treatment, alpha = 0.2)) +
  labs(title = "Urchin counts at different distances (in pie slices) from inflow",
       xlab = "Slices away from inflow",
       ylab = "Urchin slice count",
       caption = "Note: half slices are where crevices are located") +
  theme_classic()
```

```{r}
#| layout-ncol: 2
#| echo: false
#| message: false
#| label: mod-bhvr-urchdistribution

#currently, slices include half slices which are for crev, removing this for model
urch_count_non_crev <- behavior_processed %>%
  filter(slice %in% c(1,2,3,4,5,6,7,8))

urch_slice_mod <- glmmTMB::glmmTMB(urchin_slice_count ~ 
  
                        Urch_habitat_treatment + 
                        Pred_treatment + 
                        slice +
                        Urch_habitat_treatment*Pred_treatment*slice +
                        
                        (1| Trial/Tank),  
                      
                      data = urch_count_non_crev,
                      
                      family = gaussian(link = "identity"))

sjPlot::tab_model(urch_slice_mod)

predicted_urch_slice_mod <- ggpredict(model = urch_slice_mod, terms = c("Pred_treatment","Urch_habitat_treatment", "slice"))

#need to format this more - and also add urchin source hab treatment somehow.
ggplot(predicted_urch_slice_mod, aes(facet, predicted, fill = x)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), position=position_dodge()) 
```

## Key takeaways:

> Urchins graze less around a star for both kelp forest and barren treatments
>
> Grazing rates for urchins from a barren are more affected by the presence of a star than urchins from a kelp forest
